---
title: "Transatlantic Sentiment"
output: html_document
date: "2025-11-16"
---

```{r setup, include=FALSE}
library(mice)
library(naniar)
library(tidyverse)
library(rio)
library(foreign)
library(tidylog)
library(Hmisc)
library(jtools)
library(performance)  
library(see)    
library(patchwork) 
library(qqplotr)
library(plm)
library(ggplot2)
library(stargazer)
library(broom)
library(dplyr)
library(tidyr)
library(stringr)
library(lmtest)
```

##Load the Data

This loads the data into RStudio, makes a copy of the original data, and restructure the original data into a country-year level data to prepare for panel data regression with country or year fixed effects.

```{r}
setwd ("/Users/hkwang9/Library/CloudStorage/OneDrive-TheUniversityofChicago/University of Chicago/Academics/3 MACSS/Thesis/Data")
data <- import("Transatlantic Rhetoric Dataset.dta")
data0 <- data

#make a copy of the original data
ex_country <- c("Sweden", "Iceland")
data1 <- data0 %>% 
  filter(! country %in% ex_country)

#restructure the data into panel data structure 
p_data <- pdata.frame(data1, index = c("ccode", "year"))
```

##Begin the analysis

###The full model

```{r}
full <- plm(RDef ~ lag(anger, 1) + lag(fear, 1) + lag(joy, 1) + lag(sadness, 1) + lag(disgust, 1) + lag(surprise, 1), data = p_data, subset = NATO == 1, model = "pooling")
summary(full)

full_fe <- full <- plm(RDef ~ lag(anger, 1) + lag(fear, 1) + lag(joy, 1) + lag(sadness, 1) + lag(disgust, 1) + lag(surprise, 1), data = p_data, subset = NATO == 1, model = "within")
summary(full_fe)

full2 <- full <- plm(RDef ~ lag(anger, 1) + lag(fear, 1) + lag(joy, 1) + lag(sadness, 1) + lag(disgust, 1) + lag(surprise, 1) + lag(Sentiment, 1) + lag(Dem, 1) + lag(IntrastCon, 1), data = p_data, subset = NATO == 1, model = "pooling")
summary(full2)

full2_fe <- full <- plm(RDef ~ lag(anger, 1) + lag(fear, 1) + lag(joy, 1) + lag(sadness, 1) + lag(disgust, 1) + lag(surprise, 1) + lag(Sentiment, 1) + lag(Dem, 1) + lag(IntrastCon, 1), data = p_data, subset = NATO == 1, model = "within")
summary(full2_fe)

```
###The European Model 

```{r}
EU <- plm(RDef ~ lag(anger, 1) + lag(fear, 1) + lag(joy, 1) + lag(sadness, 1) + lag(disgust, 1) + lag(surprise, 1), data = p_data, subset = EU == 1, model = "pooling")
summary(EU)

EU_fe <- plm(RDef ~ lag(anger, 1) + lag(fear, 1) + lag(joy, 1) + lag(sadness, 1) + lag(disgust, 1) + lag(surprise, 1), data = p_data, subset = EU == 1, model = "within")
summary(EU_fe)

EU2 <- plm(RDef ~ lag(anger, 1) + lag(fear, 1) + lag(joy, 1) + lag(sadness, 1) + lag(disgust, 1) + lag(surprise, 1) + lag(Sentiment, 1) + lag(Dem, 1) + lag(IntrastCon, 1), data = p_data, subset = EU == 1, model = "pooling")
summary(EU2)

EU2_fe <- plm(RDef ~ lag(anger, 1) + lag(fear, 1) + lag(joy, 1) + lag(sadness, 1) + lag(disgust, 1) + lag(surprise, 1) + lag(Sentiment, 1) + lag(Dem, 1) + lag(IntrastCon, 1), data = p_data, subset = EU == 1, model = "within")
summary(EU2_fe)
```
### The Demorcrat Model 

```{r}
dem <- plm(RDef ~ lag(anger, 1) + lag(fear, 1) + lag(joy, 1) + lag(sadness, 1) + lag(disgust, 1) + lag(surprise, 1), data = p_data, subset = party == "Democrat", model = "pooling")
summary(dem)

dem_fe <- plm(RDef ~ lag(anger, 1) + lag(fear, 1) + lag(joy, 1) + lag(sadness, 1) + lag(disgust, 1) + lag(surprise, 1), data = p_data, subset = party == "Democrat", model = "within")
summary(dem_fe)

dem2 <- plm(RDef ~ lag(anger, 1) + lag(fear, 1) + lag(joy, 1) + lag(sadness, 1) + lag(disgust, 1) + lag(surprise, 1) + lag(Sentiment, 1) + lag(Dem, 1) + lag(IntrastCon, 1), data = p_data, subset = party == "Democrat", model = "pooling")
summary(dem2)


dem2_fe <- plm(RDef ~ lag(anger, 1) + lag(fear, 1) + lag(joy, 1) + lag(sadness, 1) + lag(disgust, 1) + lag(surprise, 1) + lag(Sentiment, 1) + lag(Dem, 1) + lag(IntrastCon, 1), data = p_data, subset = party == "Democrat", model = "within")
summary(dem2_fe)

```

### The Republican Model 

```{r}
gop <- plm(RDef ~ lag(anger, 1) + lag(fear, 1) + lag(joy, 1) + lag(sadness, 1) + lag(disgust, 1) + lag(surprise, 1), data = p_data, subset = party == "Republican", model = "pooling")
summary(gop)

gop_fe <- plm(RDef ~ lag(anger, 1) + lag(fear, 1) + lag(joy, 1) + lag(sadness, 1) + lag(disgust, 1) + lag(surprise, 1), data = p_data, subset = party == "Republican", model = "within")
summary(gop_fe)

gop2 <- plm(RDef ~ lag(anger, 1) + lag(fear, 1) + lag(joy, 1) + lag(sadness, 1) + lag(disgust, 1) + lag(surprise, 1) + lag(Sentiment, 1) + lag(Dem, 1) + lag(IntrastCon, 1), data = p_data, subset = party == "Republican", model = "pooling")
summary(gop2)


gop2_fe <- plm(RDef ~ lag(anger, 1) + lag(fear, 1) + lag(joy, 1) + lag(sadness, 1) + lag(disgust, 1) + lag(surprise, 1) + lag(Sentiment, 1) + lag(Dem, 1) + lag(IntrastCon, 1), data = p_data, subset = party == "Republican", model = "within")
summary(gop2_fe)

```

###sewiing the regression results together 

```{r}
# =========================================================
# 0) Shared labels / settings
# =========================================================
col_labels <- c("Full Model", "European Model", "Democrat Model", "Republican Model")

stars_cutoffs <- c(0.1, 0.05, 0.01)

note_html <- "* p < 0.1; ** p < 0.05; *** p < 0.01"
note_tex  <- "$^{*}$ p < 0.1; \\, $^{**}$ p < 0.05; \\, $^{***}$ p < 0.01"

# Labels for NO-control models (Tables 2 & 3): emotions only
cov_labels_nocontrol <- c(
  "Anger (1-year lag)",
  "Fear (1-year lag)",
  "Joy (1-year lag)",
  "Sadness (1-year lag)",
  "Disgust (1-year lag)",
  "Surprise (1-year lag)"
)

# Labels for control models (Tables 4 & 5): emotions + controls
cov_labels_control <- c(
  "Anger (1-year lag)",
  "Fear (1-year lag)",
  "Joy (1-year lag)",
  "Sadness (1-year lag)",
  "Disgust (1-year lag)",
  "Surprise (1-year lag)",
  "Sentiment (1-year lag)",
  "Liberal Democracy Index (1-year lag)",   # Dem
  "Intrastate Conflict (1-year lag)"
)

# =========================================================
# 1) Helper: drop controls from a model (for Tables 2 & 3)
#    Controls = lag(Sentiment,1) + lag(Dem,1) + lag(IntrastCon,1)
# =========================================================
drop_controls <- function(model_obj) {
  # Uses update() so you do not need to rewrite the full formula.
  # If a term isn't present, update() will just ignore it safely.
  update(
    model_obj,
    . ~ . -
      lag(Sentiment, 1) -
      lag(Dem, 1) -
      lag(IntrastCon, 1)
  )
}

# =========================================================
# 2) TEX post-processing: enforce letter-width fit + 1.0 spacing
#    Requires in LaTeX preamble:
#      \usepackage{setspace}
#      \usepackage{adjustbox}
# =========================================================
wrap_tex_letter_single <- function(tex_file) {
  x <- readLines(tex_file, warn = FALSE)

  # Insert after \centering
  x <- gsub(
    "\\\\centering",
    "\\\\centering\n\\\\setstretch{1.0}\n\\\\adjustbox{max width=\\\\textwidth}{",
    x
  )

  # Close adjustbox before \end{table}
  x <- gsub(
    "\\\\end\\{table\\}",
    "}\n\\\\end{table}",
    x
  )

  writeLines(x, tex_file)
}

# =========================================================
# 3) Writers: DOC (HTML) + TEX (LaTeX)
# =========================================================
make_doc <- function(models, outfile, cov_labels, controls_yesno, fe_yesno) {
  stargazer(
    models[[1]], models[[2]], models[[3]], models[[4]],
    type = "html",
    column.labels = col_labels,
    column.separate = c(1, 1, 1, 1),
    covariate.labels = cov_labels,
    star.cutoffs = stars_cutoffs,
    notes = note_html,
    notes.append = FALSE,
    notes.align = "l",
    add.lines = list(
      c("Controls", controls_yesno, controls_yesno, controls_yesno, controls_yesno),
      c("Country Fixed Effects", fe_yesno, fe_yesno, fe_yesno, fe_yesno)
    ),
    out = outfile
  )
}

make_tex <- function(models, outfile, cov_labels, controls_yesno, fe_yesno) {
  stargazer(
    models[[1]], models[[2]], models[[3]], models[[4]],
    type = "latex",
    column.labels = col_labels,
    column.separate = c(1, 1, 1, 1),
    covariate.labels = cov_labels,
    star.cutoffs = stars_cutoffs,
    notes = note_tex,
    notes.append = FALSE,
    notes.align = "l",
    add.lines = list(
      c("Controls", controls_yesno, controls_yesno, controls_yesno, controls_yesno),
      c("Country Fixed Effects", fe_yesno, fe_yesno, fe_yesno, fe_yesno)
    ),
    out = outfile
  )

  wrap_tex_letter_single(outfile)
}

# =========================================================
# 4) Build model sets
#    IMPORTANT CHANGE:
#      Table 2 & 3 "Full Model" will be re-estimated WITHOUT controls.
# =========================================================

# --- Tables 2 & 3: NO controls in ANY column ---
full_nc    <- drop_controls(full)
full_fe_nc <- drop_controls(full_fe)

m_table2 <- list(full_nc,    EU,    dem,    gop)      # no control, no FE
m_table3 <- list(full_fe_nc, EU_fe, dem_fe, gop_fe)  # no control, FE

# --- Tables 4 & 5: WITH controls (keep as your existing objects) ---
m_table4 <- list(full2,    EU2,    dem2,    gop2)       # control, no FE
m_table5 <- list(full2_fe, EU2_fe, dem2_fe, gop2_fe)    # control, FE

# =========================================================
# 5) Generate DOC + TEX outputs
# =========================================================

# Table 2: no control, no FE
make_doc(m_table2, "table2.doc", cov_labels_nocontrol, controls_yesno = "No",  fe_yesno = "No")
make_tex(m_table2, "table2.tex", cov_labels_nocontrol, controls_yesno = "No",  fe_yesno = "No")

# Table 3: no control, FE
make_doc(m_table3, "table3.doc", cov_labels_nocontrol, controls_yesno = "No",  fe_yesno = "Yes")
make_tex(m_table3, "table3.tex", cov_labels_nocontrol, controls_yesno = "No",  fe_yesno = "Yes")

# Table 4: controls, no FE
make_doc(m_table4, "table4.doc", cov_labels_control,   controls_yesno = "Yes", fe_yesno = "No")
make_tex(m_table4, "table4.tex", cov_labels_control,   controls_yesno = "Yes", fe_yesno = "No")

# Table 5: controls, FE
make_doc(m_table5, "table5.doc", cov_labels_control,   controls_yesno = "Yes", fe_yesno = "Yes")
make_tex(m_table5, "table5.tex", cov_labels_control,   controls_yesno = "Yes", fe_yesno = "Yes")

message("Done. Updated table2–table5 as .doc and .tex (T2/T3 have NO controls).")

# =========================================================
# 6) LaTeX preamble requirements (Overleaf)
# =========================================================
# \usepackage{setspace}
# \usepackage{adjustbox}

```


### Visualizations 
```{r}
# =========================
# A) Robust tidying for plm
# =========================
tidy_plm_robust <- function(model, model_name, group_name, vcov_type = "HC1") {
  V <- plm::vcovHC(model, type = vcov_type, cluster = "group")  # cluster by country (ccode)
  ct <- lmtest::coeftest(model, vcov. = V)

  out <- broom::tidy(ct) %>%
    rename(term = term, estimate = estimate, std.error = std.error, statistic = statistic, p.value = p.value) %>%
    mutate(
      conf.low  = estimate - 1.96 * std.error,
      conf.high = estimate + 1.96 * std.error,
      model = model_name,
      group = group_name
    )

  out
}

pretty_term <- function(x) {
  x %>%
    str_replace_all("lag\\(([^,]+), 1\\)", "\\1 (t-1)") %>%
    str_replace_all("\\bDem\\b", "Liberal Democracy Index (t-1)") %>%
    str_replace_all("\\bSentiment\\b", "Sentiment (t-1)") %>%
    str_replace_all("\\bIntrastCon\\b", "Intrastate Conflict (t-1)")
}

# =========================
# B) Build tidy dataset
# =========================
build_coef_df <- function(models_named_list, model_name) {
  # models_named_list: named list where names are groups ("Full","EU","Dem","GOP")
  bind_rows(lapply(names(models_named_list), function(g) {
    tidy_plm_robust(models_named_list[[g]], model_name = model_name, group_name = g)
  })) %>%
    filter(term != "(Intercept)") %>%
    mutate(term_pretty = pretty_term(term))
}

# =========================
# C) Coefficient plot
# =========================
plot_coefs <- function(coef_df, title) {
  ggplot(coef_df, aes(x = estimate, y = fct_reorder(term_pretty, estimate))) +
    geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.6) +
    geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
    geom_point(size = 2) +
    facet_grid(group ~ model, scales = "free_y") +
    labs(
      title = title,
      x = "Coefficient (robust 95% CI)",
      y = NULL
    ) +
    theme_minimal(base_size = 12)
}

# =========================
# D) Predicted effect plot (single predictor)
#    Uses a simple "what-if" approach:
#    set all other regressors to their means, vary one emotion.
# =========================
predict_effect <- function(model, p_data, predictor,
                           grid_n = 30, subset_filter = NULL) {

  df <- as.data.frame(p_data)

  # optionally restrict to the same subset you estimated on
  if (!is.null(subset_filter)) df <- df %>% filter(!!rlang::parse_expr(subset_filter))

  # get model frame terms from the fitted model
  f <- formula(model)
  vars_needed <- all.vars(f)

  # keep only variables used in model
  df2 <- df %>% select(any_of(vars_needed)) %>% drop_na()

  # build a "typical" row = means of numeric columns
  base_row <- df2 %>%
    summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)))

  # grid for predictor
  rng <- range(df2[[predictor]], na.rm = TRUE)
  grid <- tibble::tibble(
    val = seq(rng[1], rng[2], length.out = grid_n)
  )

  newdata <- base_row[rep(1, grid_n), , drop = FALSE]
  newdata[[predictor]] <- grid$val

  # predicted values (note: for FE models, this is "within" fitted component)
  yhat <- as.numeric(predict(model, newdata = newdata))

  tibble::tibble(
    predictor = predictor,
    x = grid$val,
    yhat = yhat
  )
}

plot_effects_4groups <- function(models_named_list, p_data, predictor, title,
                                 subset_filter = NULL) {

  df_eff <- bind_rows(lapply(names(models_named_list), function(g) {
    predict_effect(models_named_list[[g]], p_data, predictor,
                   subset_filter = subset_filter) %>%
      mutate(group = g)
  }))

  ggplot(df_eff, aes(x = x, y = yhat)) +
    geom_line() +
    facet_wrap(~ group, scales = "free_y") +
    labs(
      title = title,
      x = predictor,
      y = "Predicted RDef"
    ) +
    theme_minimal(base_size = 12)
}
```

##Alternative Explanations 

### fri 

```{r}
fri <- plm(RDef ~ lag(anger, 1) + lag(fear, 1) + lag(joy, 1) + lag(sadness, 1) + lag(disgust, 1) +  lag(surprise, 1) + lag(fri, 1), data = p_data, subset = EU == 1, model = "pooling")
summary(fri)

fri_fe <- full <- plm(RDef ~ lag(anger, 1) + lag(fear, 1) + lag(joy, 1) + lag(sadness, 1) + lag(disgust, 1) + lag(surprise, 1) + lag(fri, 1), data = p_data, subset = EU == 1, model = "within")
summary(fri)

fri2 <- full <- plm(RDef ~ lag(anger, 1) + lag(fear, 1) + lag(joy, 1) + lag(sadness, 1) + lag(disgust, 1) + lag(surprise, 1) + lag(fri, 1) + lag(Sentiment, 1) + lag(Dem, 1) + lag(IntrastCon, 1), data = p_data, subset = EU == 1, model = "pooling")
summary(fri2)

fri2_fe <- full <- plm(RDef ~ lag(anger, 1) + lag(fear, 1) + lag(joy, 1) + lag(sadness, 1) + lag(disgust, 1) + lag(surprise, 1) + lag(fri, 1) + lag(Sentiment, 1) + lag(Dem, 1) + lag(IntrastCon, 1), data = p_data, subset = EU == 1, model = "within")
summary(fri2_fe)
```

### proximity 

```{r}
proxy_raw <- plm(RDef ~ lag(anger, 1) + lag(fear, 1) + lag(joy, 1) + lag(sadness, 1) + lag(disgust, 1) + lag(surprise, 1) + log_proxy + lag(Sentiment, 1) + lag(Dem, 1) + lag(IntrastCon, 1), data = p_data, subset = NATO == 1, model = "pooling")
summary(proxy_raw)

proxy_EU <- plm(RDef ~ lag(anger, 1) + lag(fear, 1) + lag(joy, 1) + lag(sadness, 1) + lag(disgust, 1) + lag(surprise, 1) + log_proxy + lag(Sentiment, 1) + lag(Dem, 1) + lag(IntrastCon, 1), data = p_data, subset = EU == 1, model = "pooling")
summary(proxy_EU)

proxy_Dem <- plm(RDef ~ lag(anger, 1) + lag(fear, 1) + lag(joy, 1) + lag(sadness, 1) + lag(disgust, 1) + lag(surprise, 1) + log_proxy + lag(Sentiment, 1) + lag(Dem, 1) + lag(IntrastCon, 1), data = p_data, subset = party == "Democrat", model = "pooling")
summary(proxy_Dem)

proxy_GOP <- plm(RDef ~ lag(anger, 1) + lag(fear, 1) + lag(joy, 1) + lag(sadness, 1) + lag(disgust, 1) + lag(surprise, 1) + log_proxy + lag(Sentiment, 1) + lag(Dem, 1) + lag(IntrastCon, 1), data = p_data, subset = party == "Republican", model = "pooling")
summary(proxy_GOP)

```





